---
Title: 'Reproducible Research: Project 1'
---

### Loading and preprocessing the data

Load the data:
```{r loaddata, echo=TRUE, cache=TRUE}
data <- read.csv("./data/activity.csv", header=TRUE)
```

Show the original data(first few rows):
```{r showdata, echo=TRUE, cache=TRUE}
head(data)
```


### What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

Calculate the total number of steps based on each days
```{r stepsDay, echo=TRUE, cache=TRUE}
stepsDay <- aggregate(steps~date, data, sum)
head(stepsDay)
```

Plot a histogram of the total number of steps taken each day
```{r histogram1, echo=TRUE, cache=TRUE}
#Colour taken from Hex color code picker
hist(stepsDay$steps, col= "#42f4bc", xlab = "Steps", main = "Total steps taken per day")
```

Calculate the mean and median of total steps taken
```{r mean1, echo=TRUE, cache=TRUE}
mean1 <- mean(stepsDay$steps)
mean1
median1 <- median(stepsDay$steps)
median1
```
### What is the average daily activity pattern?
Subset the original data to calculate the average number of steps taken of the 5-minute interval
```{r avgSteps, echo=TRUE, cache=TRUE}
avgSteps <- aggregate(steps~interval, data, mean)
head(avgSteps)
```

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r plot1, echo=TRUE, cache=TRUE}
with(avgSteps, plot(interval, steps
                , type="l", main="Average steps taken in 5-min interval"
                , col = "#41b8f4")) 
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

The max number of steps & corresponding interval are:
```{r intervalmax, echo=TRUE, cache=TRUE}
max <- max(avgSteps$steps)
max

subsetData <- subset(avgSteps, steps == max)
subsetData$interval
```

## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r missing, echo=TRUE}
colSums(is.na(data))
```

Create a new column "newSteps" contains the average steps
```{r newsteps, echo=TRUE, cache=TRUE}
data$newSteps <- avgSteps$steps
head(data)
```

Fill in the missing data in the original steps column
```{r fill, echo=TRUE, cache=TRUE}
data$steps[is.na(data$steps)] <- data$newSteps
head(data)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r newdata, echo=TRUE, cache=TRUE}
data$newSteps <- NULL
newdata <- data
head(newdata)
```

Subset in the new data set to calculate the total number of steps per day
```{r sumdata, echo=TRUE, cache=TRUE}
sumdata <- aggregate(steps~date, newdata, sum)
head(sumdata)
```

Make a histogram of the total number of steps taken each day
```{r histogram2, echo=TRUE, cache=TRUE}
hist(sumdata$steps, col = "#41dff4"
     , xlab = "Steps"
     , main = "Total steps per day (with missing values filled in)")
```

Calculate the mean and median of the total number of steps taken per day
```{r mean2, echo=TRUE, cache=TRUE}
mean2 <- mean(sumdata$steps)
mean2
median2 <- median(sumdata$steps)
median2
```

Calculate the differences of the mean and median between the first(with missing values) and second part(with missing filled in)
```{r diff, echo=TRUE}
mean2 - mean1
median2 - median1
```
## Are there differences in activity patterns between weekdays and weekends?
Create two new variables weekdays & weekends as factor variables.

```{r check, echo=TRUE, cache=TRUE}
library(timeDate)
newdata$Weekday <- isWeekday(newdata$date)
head(newdata)
```

calculate the average steps for weekday and weekend
Weekday
```{r weekday, echo=TRUE, cache=TRUE}
weekday <- subset(newdata, newdata$Weekday == "TRUE")
weekdayMean <- aggregate(steps ~ interval, data = weekday, mean)
head(weekdayMean)
```

Weekend
```{r weekend, echo=TRUE, cache=TRUE}
weekend <- subset(newdata, newdata$Weekday == "FALSE")
weekendMean <- aggregate(steps ~ interval, data = weekend, mean)
head(weekendMean)
```

Make the panel plot to calculate the average number of steps taken for weekday and weekend
```{r panel, echo=TRUE, cache=TRUE}
par(mfrow=c(1,2), mar=c(5,5,2,2))
plot(weekdayMean$interval, weekdayMean$steps, ylim=c(1, 250)
     , xlab = "interval", ylab = "Number of steps"
     , main ="Weekday", col ="blue", type="l")

plot(weekendMean$interval, weekendMean$steps, ylim=c(1,250)
     , xlab = "interval", ylab = "Number of steps"
     , main ="Weekend", col ="red", type="l")

```
